Viscous Transport in Eroding Porous Media

### 2019-09-24 docs/transport/erosion_transport.tex
NM

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Porous media flow plays an important role in many environmental and
industrial applications.  Depending on the application, length scales
can vary from $10^{-6}\mathrm{m}$ to
$10^{-1}\mathrm{m}$~\cite{mil-chr-imh-mcb-ped1998}, velocity scales can
be as small as $10^{-1}\mathrm{m/day}$~\cite{kut-scr-dav-ham1995}, \nick{and
the scales of a single porous media can range over several orders of
magnitude.}\todo[inline]{Nick: I am confused by the last part of the sentence. What 'scale' are you referring to? Length? You already described the length scales in the first part of the sentence. Maybe you mean to distinguish between the grain length scales and the overall length scales of the system. If so, you can use adjectives like `grain' and `macroscopic' or `overall' to clarify that.} 
Numerical methods that accurately address all of these scales
are essential to develop meaningful constitutive relationships between
the microscopic and macroscopic scales~\cite{mil-chr-imh-mcb-ped1998},
to characterize dispersion~\cite{saf1959}, and to quantify
mixing~\cite{leb-den-dav-bol-car-dec-bou2011, den-leb-eng-bij2011}.
Examples of coarse-grained models for porous media flow include
pressure-saturation-permeability
relationships~\cite{mil-chr-imh-mcb-ped1998}, permeability-porosity
relationships~\cite{dar-mcc1998, car1937, won-kop-tom1984}, first
passage time distributions~\cite{ber-sch-sil2000, hym-den-hag-kan2019,
cve-che-wen1996}, tortuosities~\cite{hak-com-den2019, mat-kha-koz2008,
dud-koz-mat2011, kop-kat-tim1996}, concentration
distributions~\cite{ica-den2019, bel-sal-rin1992}, flow
distributions~\cite{ali-par-wei-bre2017}, geometry
connectivity~\cite{knu-car2005, wes-blo-gra2001}, pore
distributions~\cite{ali-par-wei-bre2017}, anomalous
dispersion~\cite{dea-qua-bir-jua2018, den-cor-sch-ber2004,
sie-ili-pri-riv-gua2019}, and more.

Flow in porous media is further complicated when there is a two-way
coupling between the flow and the geometry. For example, the grains of a
porous media evolve in time in applications involving
melting~\cite{bec-vis1998}, dissolution~\cite{kan-zha-che-he2002},
deposition~\cite{joh-eli1995}, and biofilm
growth~\cite{tan-val-wer2015}.  We focus on erosion, a fluid-mechanical
process that is prevalent in many geophysical and hydrological
applications.  When a porous geometry erodes, certain qualitative
characteristics are unveiled that affect transport through the geometry.
For example, an eroded geometry often contains channels of high porosity
that connect the inlet to the outlet, and this results in velocities
that vary over several orders of magnitude~\cite{all-hea-lab-rei2002}.
Moreover, channelized geometries are heterogeneous and anisotropic, and
this affects the transport of quantities such as
contaminants~\cite{cve-che-wen1996, dag1987, kon-bre1978} and
heat~\cite{nil-sto1990, ree-sto1995}.  To characterize transport in
eroded geometries, we complete two undertakings: forming realistic
eroded porous geometries and characterizing transport through these
geometries.  The modelling effort and numerical methods used to simulate
erosion are largely based on our previous work~\cite{qua-moo2018}.
However, we improve our numerical methods so that simulations of more
realistic dense suspensions of eroding grains are possible
(Figures~\ref{fig:Eroding50vel} and~\ref{fig:Eroding50vort}).  To
characterize transport in eroded porous geometries, we apply statistical
methods to develop coarse-grained variables.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Porous media flow plays an important role in many environmental and
industrial applications.  Depending on the application, length scales
can vary from $10^{-6}\mathrm{m}$ to
$10^{-1}\mathrm{m}$~\cite{mil-chr-imh-mcb-ped1998}, velocity scales can
be as small as $10^{-1}\mathrm{m/day}$~\cite{kut-scr-dav-ham1995}, \nick{and
the scales of a single porous media can range over several orders of
magnitude.}
%
\todo[inline]{Nick: I am confused by the last part of the sentence (highlighted in red). What `scale' are you referring to? Length? You already described the length scales in the first part of the sentence. Maybe you mean to distinguish between the grain length scales and the overall length scales of the system. If so, you can use adjectives like `grain' and `macroscopic' or `overall' to clarify that.} 
%
Numerical methods that \st{accurately address all of these} \nick{resolve this range of} scales offer the ability to: (i) characterize dispersion~\cite{saf1959}, (ii) quantify mixing~\cite{leb-den-dav-bol-car-dec-bou2011, den-leb-eng-bij2011}, and (iii) potentially develop meaningful constitutive relationships that link the microscopic and macroscopic realms~\cite{mil-chr-imh-mcb-ped1998}. 
%
\todo[inline]{Nick: I reworded the above sentence. See if you agree with my changes.}
%
Examples of coarse-grained models for porous media flow include
pressure-saturation-permeability
relationships~\cite{mil-chr-imh-mcb-ped1998}, permeability-porosity
relationships~\cite{dar-mcc1998, car1937, won-kop-tom1984}, first
passage time distributions~\cite{ber-sch-sil2000, hym-den-hag-kan2019,
cve-che-wen1996}, tortuosities~\cite{hak-com-den2019, mat-kha-koz2008,
dud-koz-mat2011, kop-kat-tim1996}, concentration
distributions~\cite{ica-den2019, bel-sal-rin1992}, flow
distributions~\cite{ali-par-wei-bre2017}, geometry
connectivity~\cite{knu-car2005, wes-blo-gra2001}, pore
distributions~\cite{ali-par-wei-bre2017}, anomalous
dispersion~\cite{dea-qua-bir-jua2018, den-cor-sch-ber2004,
sie-ili-pri-riv-gua2019}, and more.

Flow in porous media is further complicated when there is a two-way
coupling between the flow and the geometry. For example, the grains of a
porous media evolve in time in applications involving
melting~\cite{bec-vis1998, rycroft2016asymmetric, morrow2019moving}, dissolution~\cite{kan-zha-che-he2002, mac2015shape, wykes2018self},
deposition~\cite{joh-eli1995}, biofilm growth~\cite{tan-val-wer2015}, and \nick{crack formation} \cite{cho2019crack}.  
%
\todo[inline]{Nick: I added some references. Check out the last one on crack formation in shrinkable, granular packings. It is really cool work done by an excellent experimentalist Sujit Datta. I met him and he voiced a strong interest in our work, so wanted to make you aware.}
%
We focus on erosion, a fluid-mechanical
process that is prevalent in many geophysical and hydrological
applications.  When a porous geometry erodes, certain qualitative
characteristics are unveiled that affect transport through the geometry.
For example, an eroded geometry often contains channels of high porosity
that connect the inlet to the outlet, and this results in velocities
that vary over several orders of magnitude~\cite{all-hea-lab-rei2002}.
Moreover, channelized geometries are heterogeneous and anisotropic, and
this affects the transport of quantities such as
contaminants~\cite{cve-che-wen1996, dag1987, kon-bre1978} and
heat~\cite{nil-sto1990, ree-sto1995}.  To characterize transport in
eroded geometries, we complete two undertakings: forming realistic
eroded porous geometries and characterizing transport through these
geometries.  The modelling effort and numerical methods used to simulate
erosion are largely based on our previous work~\cite{qua-moo2018}.
However, we improve our numerical methods so that simulations of more
realistic dense suspensions of eroding grains are possible
(Figures~\ref{fig:Eroding50vel} and~\ref{fig:Eroding50vort}).  To
characterize transport in eroded porous geometries, we apply statistical
methods to develop coarse-grained variables.
